Hardware-accelerated Rendering of Intersecting Volumes using Boolean Operations

### 2013-04-21 paper.tex
Broke up introduction in second approach
Moved notes about characteristics of second approach to end of subsection
Summary of second approach in its introduction
Fully-detailed second approach

The second approach uses a Boolean AND operation together with a Boolean XOR.
It is mostly CPU-based, since it requires both Boolean operations to be done on
the CPU.  The extra work on the CPU results in less rendering passes, however.
The pseudocode for the approach is shown below.

@@

>>>>

The second approach uses a Boolean AND operation together with a Boolean XOR.
Here the idea is to break up both volumes into pieces around the intersection.
All the pieces behind the intersection are rendered first, then the intersection
itself, and then all the pieces in front of the intersection.

@@

Like the first approach, the first step is to set up the textures.  There are
two textures for the volumes, two to store the texture coordinates of the back
faces, and one to hold results between passes.  The textures to store the
texture coordinates and results between passes will be used with external
framebuffer objects, and should thus be at least as big as the window.

\begin{Verbatim}[fontsize=\small]
  <texture id="volume-texture-1" file="volume-1.vlb" />
  <texture id="volume-texture-2" file="volume-2.vlb" />
  <texture id="coords-texture-1" size="768" />
  <texture id="coords-texture-2" size="768" />
  <texture id="results-texture" size="768" />
\end{Verbatim}

Storage should also be created to act as a depth buffer for some passes.  Since
we do not need to read from it, it can just be a \emph{renderbuffer}.  It also
needs to be at least as big as the window.

\begin{Verbatim}[fontsize=\small]
  <renderbuffer id="depth-renderbuffer" format="depth" size="768" />
\end{Verbatim}

Next the results texture should be cleared.

\begin{Verbatim}[fontsize=\small]
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <clear color="0 0 0 0" />
  </framebuffer>
\end{Verbatim}

Then the pieces behind the intersection should be used to render the volumes to
the results texture.  Because there may be multiple pieces behind the
intersection and we only want the texture coordinates of the faces that are
farthest back, we first clear the depth buffer to zero and set the depth
function to \emph{GREATER}.  Once we have those coordinates stored, we do the
opposite by clearing the depth buffer to one and setting the depth function back
to \emph{LESS}.  That way we render using the faces that are farthest forward.
Note in the code snippet below we only show the first volume.

\begin{Verbatim}[fontsize=\small]
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <attachment usage="depth" link="depth-renderbuffer" />
    <clear depth="0.0" />
    <depth function="greater" />
    <cull faces="front" />
    <use program="primary-coords-program">
      <group id="pieces-of-cube-1-behind-intersection-group">
        <booleanXor of="cube-1 cube-2" hide="cube-2" filter="gequal" />
      </group>
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <attachment usage="depth" link="depth-renderbuffer" />
    <clear depth="1.0" />
    <depth function="less" />
    <cull faces="back" />
    <use program="first-pass-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <uniform name="BackFacesTexture" link="coords-texture-1" />
      <uniform name="VolumeTexture" link="volume-texture-1" />
      <instance of="pieces-of-cube-1-behind-intersection-group" />
    </use>
  </framebuffer>
\end{Verbatim}

Following our pattern, we now render the intersection itself.  Notice we do not
need a depth buffer for this pass, although we do need to make sure both sets of
texture coordinates are stored and are available when we render the
intersection.

\begin{Verbatim}[fontsize=\small]
  <cull faces="front" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <group id="intersection-group">
        <booleanAnd of="cube-1 cube-2" />
      </group>
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="coords-texture-2" />
    <use program="secondary-coords-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <instance of="intersection-group" />
    </use>
  </framebuffer>
  <cull faces="back" />
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <use program="second-pass-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <uniform name="ResultsTexture" link="results-texture" />
      <uniform name="CoordsTexture1" link="coords-texture-1" />
      <uniform name="CoordsTexture2" link="coords-texture-2" />
      <uniofrm name="VolumeTexture1" link="volume-texture-1" />
      <uniofrm name="VolumeTexture2" link="volume-texture-2" />
      <instance of="intersection-group" />
    </use>
  </framebuffer>
\end{Verbatim}

Next we render any pieces of the first cube in front of the intersection to the
results texture.  Similar to the first pass, we use the depth buffer to make
sure we only take the texture coordinates of the faces we want.  Again we omit
the code to render the second cube.

\begin{Verbatim}[fontsize=\small]
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <attachment usage="depth" link="depth-renderbuffer" />
    <cull faces="front" />
    <depth function="greater" />
    <clear depth="0.0" />
    <use program="primary-coords-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <group id="pieces-of-cube-1-in-front-of-intersection">
        <booleanXor of="cube-1 cube-2" hide="cube-2" filter="less" />
      </group>
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <attachment usage="depth" link="depth-renderbuffer" />
    <depth function="less" />
    <cull faces="back" />
    <use program="third-pass-program">
      <uniform name="MVPMatrix" usage="viewprojection" />
      <uniform name="ResultsTexture" link="results-texture" />
      <uniform name="CoordsTexture" link="coords-texture-1" />
      <uniform name="VolumeTexture" link="volume-texture-1" />
      <insance of="pieces-of-cube-1-in-front-of-intersection" />
    </use>
  </framebuffer>
\end{Verbatim}

Finally we render the results to the screen by drawing both cubes again.  The
shader program for this final pass just outputs whatever is in the results
texture.  We need this final pass because depending on how the scene is laid out
there may not be any pieces behind or in front of the intersection, meaning the
first and third passes may not be run.

\begin{Verbatim}[fontsize=\small]
  <use program="final-pass-program">
    <uniform name="ResultsTexture" link="results-texture" />
    <translate x="-1.0">
      <uniform name="MVPMatrix" usage="modelviewprojection" />
      <cube id="cube-1" />
    </translate>
    <translate x="+1.0">
      <uniform name="MVPMatrix" usage="modelviewprojection" />
      <cube id="cube-2" />
    </translate>
  </use>
\end{Verbatim}

It is mostly CPU-based, since it requires both Boolean operations to be done on
the CPU.  The extra work on the CPU results in less rendering passes, however.
The pseudocode for the approach is shown below.
